---
title: "dimReduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dimReduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Overview
This vignette describes the usage of the `dimReduction` parameter in the `pandemonium` app and how new functions can be written for it.


## what the function takes
mat = n x m matrix of coordinates to reduce
dist = n x n distance matrix.

## What the function should return:
`dimReduction(mat = mat, ist = dist)$Y` = n x 2 matrix or data frame with dimension reduced coordinates


# How to write new functions
The standard way to write a new function to be passed to `dimReduction` is as a wrapper for a common dimension reduction algorithm.
The defined function must be able to take `mat` and `dist` as parameters. If only one of them is used it is best to write the function with space for additional parameters with `...` as seen in the following example.
```{r,echo=TRUE,eval=FALSE}
tSNE_mat <- function(mat,...) {
  coord_red = list()
  coord_red$Y = Rtsne::Rtsne(mat)$Y
  coord_red
}

tSNE_dist <- function(dist,...) {
  coord_red = list()
  coord_red$Y = Rtsne::Rtsne(dist, is.dist = TRUE)$Y
  coord_red
}

newPandemonium(df, dim_reduction = list(tSNE_mat = tSNE_mat, tSNE_dist = tSNE_dist))
```

Some algorithms may take a dist object instead of a distance matrix. This can be done by converting the distance matrix using `stats::as.dist()`. As done with `umap`. 

```{r,echo=TRUE,eval=FALSE}
umap <- function(dist,...) {
  ret <- list()
  ret$Y <- uwot::umap(stats::as.dist(dist))
  ret
}

newPandemonium(df, dim_reduction = list(umap = umap))
```
